Population Inference of Eccentric Binary Black Holes

### 2023-05-09 paper/paper.tex
Adding Power law subsection and VT Subsection

\begin{itemize}
    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
    \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
    $p(\Lambda|d_1,d_2,d_3,...,d_N)= \frac{p(\Lambda)*p(d_1,d_2,d_3,...,d_N|\Lambda)}{p(d_1,d_2,d_3,...,d_N)}$
    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
    \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$. 
    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
    $\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)$
    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$. $l_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
    \item now we can find the posterior by Baye's theorem. \\
    $p(\mathcal{R,\Lambda})\propto p(\mathcal{R},\Lambda) * \mathcal{L}(\mathcal{R},\Lambda)$
    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.

\subsection{Power law Model}

@@

>>>>

\begin{itemize}
    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
    \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
\begin{equation}
\label{eq:Bayes}    
p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)}
\end{equation}

    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
    \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$. 
    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
\begin{equation}
\label{eq: likelihood}
\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)    
\end{equation}

    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$. $l_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
    \item now we can find the posterior by Baye's theorem. \\
    $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$
    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.

\end{itemize}
\subsection{VT Estimation}
\begin{itemize}
    \item Estimation of VT (sensitivity of the LVK) only depend on the mass. For the sake of simplicity {\color{red}(physics reason)}, we consider only non-spinning, non-eccentric and nonprecessing BH Binaries. 
    \item Sources are detected if SNR is greater than 8.
    \item V is characteristic volume in which source can be detected by LVK. It is orientation averaged sensitive 3-volume.
\begin{equation}
\label{eq:volume}
V(\lambda) = \int P((<D(z))/D_n(\lambda))\frac{dV_c}{dz}\frac{dz}{1+z}
\end{equation}    
    \item where $D(z)$ is the luminosity distance for redshift z, $V_c$ is the comoving volume.

    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this senstivity for the time $T$, the average number of detections would be as follows: \\
\begin{equation}
\label{eq:mu}
  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda  
\end{equation}
    \item where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic paramter $\lambda$. Keep in mind that $\lambda$ is equal to all the 15 parameters of a binary.
\end{itemize}

\subsection{Power law Model}
The following is the power law model. We added the Gaussian eccentricity in it.

\begin{align}
\label{eq:plawg}
p(m_1,m_2,e) = &C(\alpha,k_m,m_{min},m_{max},M_{max},e)  
\nonumber \\ & \frac{(\frac{m_2}{m_1})^{k_m}m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
\nonumber \\ &
\sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
\end{align}
We use $k_m=0$, because GW networks are much more senstive to more massive BHs with $M\geq 200 M_\odot$

@@

The probability density would be non zero only for the two following conditions.

\begin{itemize}
    \item $m_{min}\leq m_2 \leq m_1 \leq m_{max}$
    \item $m_1+m_2 \leq M_{max}$
\end{itemize}
Where $M_{max}$ depends on the detector (LVK)
